### 2025-01-14 index.qmd
Testing embedded chem structures

---
title: "The Lipid Interactome Repository"
format: 
  html:
    embed-resources: false
standalone: false
---

A blurb about the objective of this site, maybe a gallery of images from select publications to put here.

```{=html}
<div style="text-align: center; margin: 20px;">
    <a href="LipidProbe/PhosphatidicAcid.qmd" style="text-decoration: none;">
        <button style="background-color: #3498db; color: white; font-size: 20px; padding: 15px 30px; border: none; border-radius: 8px; margin: 10px; cursor: pointer;">
            Phosphatidic Acid
        </button>
    </a>
    <a href="LipidProbe/Phosphatidylethanolamine.qmd" style="text-decoration: none;">
        <button style="background-color: #e74c3c; color: white; font-size: 20px; padding: 15px 30px; border: none; border-radius: 8px; margin: 10px; cursor: pointer;">
            Phosphatidylethanolamine
        </button>
    </a>
		<br>
		<a href="LipidProbe/Diacylglycerol.qmd" style="text-decoration: none;">
        <button style="background-color: #3277a8; color: white; font-size: 20px; padding: 15px 30px; border: none; border-radius: 8px; margin: 10px; cursor: pointer;">
            Diacylglycerol
        </button>
    </a>
		<br>
    <a href="LipidProbe/Sphinganine.qmd" style="text-decoration: none;">
        <button style="background-color: #2ecc71; color: white; font-size: 20px; padding: 15px 30px; border: none; border-radius: 8px; margin: 10px; cursor: pointer;">
            Sphinganine
        </button>
    </a>
		<a href="LipidProbe/Sphingosine.qmd" style="text-decoration: none;">
        <button style="background-color: #eba434; color: white; font-size: 20px; padding: 15px 30px; border: none; border-radius: 8px; margin: 10px; cursor: pointer;">
            Sphingosine
        </button>
    </a>

	<a href="LipidProbe/8-3_FattyAcid.qmd" style="text-decoration: none;">
        <button style="background-color: #ba34eb; color: white; font-size: 20px; padding: 15px 30px; border: none; border-radius: 8px; margin: 10px; cursor: pointer;">
            8-3 Fatty Acid
        </button>
    </a>

    <a href="LipidProbe/1-10_FattyAcid.qmd" style="text-decoration: none;">
        <button style="background-color: #a83260; color: white; font-size: 20px; padding: 15px 30px; border: none; border-radius: 8px; margin: 10px; cursor: pointer;">
            1-10 Fatty Acid
        </button>
    </a>
</div>
```