### 2025-01-16 IndividualStudies/RM_2021.qmd
Updating individual study pages

---
title: "Muller et al. 2021"
subtitle: "Synthesis and Cellular Labeling of Multifunctional Phosphatidylinositol *Bis*- and *Tris*phosphate Derivatives"
format: 
  html:
    embed-resources: false
standalone: false
---

```{r setup, echo=FALSE, include=FALSE}
## intitialize libraries
library(dplyr)
library(tidyr)
library(plotly)
library(ggplot2)
library(grid)
library(readr)
library(rmarkdown)
library(emojifont)
library(gt)
library(crosstalk)
library(DT)
library(htmltools)

## Opening the source file for the ggplot custom theme
source("ggplot_styles.R")

# Opening the datatable for this page WHILE TESTING
# RM_2021_data <- readr::read_csv("IndividualStudies/DataTables/RM_Angewandte_2021_download.csv")

# Opening the datatable for this page WHILE RENDERING
RM_2021_data <- readr::read_csv("DataTables/RM_Angewandte_2021_download.csv")

```

::: {.callout-note collapse="true"}

## Study details

#### Authors
Rainer Muller, Ana Kojic, Mevlut Citir, Carsten Schultz

#### Journal
Angewandte Chemie

<https://doi.org/10.1002/anie.202103599>{target="_blank"}

#### Abstract

We synthesized the first multifunctionalized phosphoinositide polyphosphate derivatives featuring a photo-removable protecting group (“cage”), a photo-crosslinkable diazirine group, and a terminal alkyne group useful for click chemistry. We demonstrate that the lipid derivatives readily enter cells. After photo-crosslinking, cell fixation and fluorescent tagging via click chemistry, we determined the intracellular location of the lipid derivatives before and after uncaging of the lipids. We find that there is rapid trafficking of PI(3,4)P2 and PI(3,4,5)P3 derivatives to the plasma membrane, opening the intriguing possibility that there is active transport of these lipids involved. We employed the photo-crosslinking and click chemistry functions to analyze the proteome of PI(3,4,5)P3-binding proteins. From the latter, we validated by RNAi that the putative lipid binding proteins ATP11A and MPP6 are involved in the transport of PI(3,4,5)P3 to the plasma membrane.

#### Lipid probes utilized

[PI(3,4,5)P~3~](/LipidProbe/PhosphatidylinositolTrisphosphate.qmd)

#### Cell line analyzed

[HeLa](https://www.atcc.org/products/ccl-2){target="_blank"}

#### Mass spectrometry quantification method

Peptide spectral counting

#### Additional sample preparation <span class="help-symbol" title="e.g. sub-cellular fractionation, depletion, etc.">?</span>
:::

## Data visualization

#### Heat map

#### Venn Diagram

## Data download

{{< downloadthis DataTables/RM_Angewandte_2021_download.csv dname="RM_Angewandte_2021">}}