### 2025-01-24 TrifunctionalLipidProbesOverview.qmd
small update to overview text

---
title: "Multifunctional Lipid Probes Background"
author: "Gaelen Guzman"
format: 
  html:
    embed-resources: false
standalone: true
---

Functionalized lipid probes are at the intersection of several niche biotechnologies. This overview will serve as a brief summary of the capabilities, nuances, and caveats of their utility. Additional resources and published review articles will be referenced throughout and linked below for greater detail.

## Introduction: Why study lipid interactomes?

## Why are functional modifications necessary in studying lipid interactions?

Synthetic lipid analogs have greatly enabled the exploration of lipid biology. However, it may not be obvious why such analogs are such valuable tools.

While, genetic editing has enabled vast advances in the study of nucleic acids and proteins, there has been no obvious parallel for studying lipids. Using CRISPR or *de novo* gene synthesis, one can subtly alter a sequence to enable visualization and pulldown without greatly affecting the structure or function of the target gene. To name a few examples, one can: mutate the active site of a protein to change substrate specificity, swap in a different subcellular localization sequence to translocate the target protein to a different organelle, swap domains to alter protein::protein binding, fuse the target protein to any number of fluorescent domains to visualize a protein's localization, or add an affinity handle to enable the enrichment of the target protein and any other protein interactors [REF].

Lipids, however, are an entirely different story altogether. Foremost of these differences is that lipids are the product of enzymatic synthesis and thus their structure cannot be directly modified within a living biological system. For much of the history of lipid research, studies utilized whole gene knockout or knock-in paradigms to alter the global lipid composition of the cell. Through these methods, we know much about the localization and broad behavior of lipid classes -- but the biological roles of select lipid species are historically challenging to explore [REF].

Further, lipids are significantly smaller than even the smallest protein -- composed of dozens of atoms, as opposed to hundreds or thousands -- and, thus, their biophysical properties are much more sensitive to modification. It is nearly impossible to fuse a lipid with a fluorescent protein (such as GFP) or classical affinity handle (such as an HA tag) without drastically affecting the biological activity of the lipid [REF].

Finally, each of the tens of thousands of distinct lipid structures is a part of a rapidly shifting metabolic network: any given lipid can be cleaved, modified, or converted into other lipid species within short time frames. Some lipids, such as sphingosine have half-lives in the minutes [REF].

For these reasons, synthetic analogs to native lipids have been a great utility toward probing the biological function of select lipid species. The chief modifications described in this repository

:::{.border}
```{r setup, echo=FALSE, include=FALSE}
library(plotly)

library(plotly)

# Load your image
image_url <- "LipidProbe/Structures/SphingosineStructure.png"

# Define hover regions (as rectangles)
shapes <- list(
  list(type = "rect", x0 = 0, x1 = 32, y0 = 50, y1 = 85,
       line = list(color = "red", width = 2),
       fillcolor = "rgba(255, 0, 0, 0.1)",  # Transparent fill
       name = "Region 1", hoverinfo = "none"),
  list(type = "rect", x0 = 50, x1 = 70, y0 = 40, y1 = 60,
       line = list(color = "blue", width = 2),
       fillcolor = "rgba(0, 0, 255, 0.1)",  # Transparent fill
       name = "Region 2", hoverinfo = "none")
)

# Define hover annotations
hover_annotations <- data.frame(
  x = c(20, 60),  # Midpoints of rectangles
  y = c(80, 50),
  text = c("Region 1 Info", "Region 2 Info")
)

# Create the plot
plot <- plot_ly() %>%
  add_trace(
    data = hover_annotations,
    x = ~x, y = ~y, text = ~text,
    mode = "markers",
    marker = list(opacity = 0),  # Invisible markers
    hoverinfo = "text"  # Show hover text
  ) %>%
  layout(
    shapes = shapes,
    images = list(
      list(
        source = image_url,
        xref = "x",
        yref = "y",
        x = 0,
        y = 100,
        sizex = 100,
        sizey = 100,
        xanchor = "left",
        yanchor = "top",
        layer = "below"
      )
    ),
    xaxis = list(visible = FALSE, range = c(0, 100)),
    yaxis = list(visible = FALSE, range = c(0, 100))
  )
```

```{r, echo=FALSE, warnings=FALSE}
# Render the plot
plot

```
:::

:::{.border}
```{=html}
<div style="position: relative; display: inline-block;">
  <img src="LipidProbe/Structures/SphingosineStructure.png" alt="Interactive Image" style="width: 500px;">

  <!-- Popup 1 -->
  <div style="position: absolute; top: 20%; left: 10%; cursor: pointer;">
    <span class="popup-hover" title="Popup for Point 1">🔴</span>
  </div>

  <!-- Popup 2 -->
  <div style="position: absolute; top: 70%; left: 50%; cursor: pointer;">
    <span class="popup-hover" title="Popup for Point 2">🔵</span>
  </div>

  <!-- Hover region 1 -->
  <div style="position: absolute; top: 20%; left: 10%; width: 15%; height: 15%; border: 2px solid red; box-sizing: border-box;"
       title="Region 1 Info"></div>

  <!-- Hover region 2 -->
  <div style="position: absolute; top: 50%; left: 50%; width: 20%; height: 20%; border: 2px solid blue; box-sizing: border-box;"
       title="Region 2 Info"></div>
</div>
</div>

<style>
  .popup-hover:hover::after {
    content: attr(title);
    position: absolute;
    top: -30px;
    left: 0;
    background: white;
    color: black;
    padding: 5px;
    border: 1px solid #ccc;
    border-radius: 5px;
    white-space: nowrap;
    box-shadow: 0px 4px 6px rgba(0, 0, 0, 0.1);
  }
</style>
```
:::

### Terminal Alkyne

### Diazirine: A photo-activated crosslinker

### Coumarin Cage: A photo-activated protection moiety

## *P*hoto*a*ctivatable & *C*lickable Probes

## Trifunctional Lipid Probes

## Experimental schema

### Cell treatment

### Protein collection

### Mass spectrometry analysis

#### Spectral counting

#### Isobaric mass tagging

:::{.border}
![](SiteFigures/BaselineProteomics_ExperimentalDesign.png){group="Overview_gallery"}
:::

\br

(Here's an example of what references would look like on the site: [@thomasTrifunctionalLipidDerivatives2025; @schultzFlashClickMultifunctionalized2022a; @schultzChemicalToolsLipid2023])