### 2025-01-28 LipidProbe/EnrichedHitsComparison.qmd
Added a shiny app

---
title: "Comparing the Enriched Hits of Different Probes"
---

<!-- Data wrangling, will comment out most of this when it's all done. 
Pulling the enriched hits from each probe, then making a long dataframe that can be plopped into a Venn diagram adjacent visualization -->
```{r setup, echo=FALSE,include=FALSE}
library(dplyr)
library(tidyr)
library(readr)

# # Making a big shared dataset with all of the TMT-quantified lipid probes - this is a little hand-wavey because some were run at different times... TBD if we want to keep this.
# PA_data <- read_csv("LipidProbe/DataSets/PA_HeLa_AT_2025.csv") |>
# 	select(LipidProbe, gene_name, logFC, pvalue, hit_annotation)

# PE_data <- read_csv("LipidProbe/DataSets/PE_HeLa_AT_2025.csv") |>
# 	select(LipidProbe, gene_name, logFC, pvalue, hit_annotation)

# Spa_data <- read_csv("LipidProbe/DataSets/Spa_Huh7_SF_2024.csv") |>
# 	select(LipidProbe, gene_name, logFC, pvalue, hit_annotation)

# Sph_data <- read_csv("LipidProbe/DataSets/Sph_Huh7_SF_2024.csv") |>
# 	select(LipidProbe, gene_name, logFC, pvalue, hit_annotation)

# FA_8_3_data <- read_csv("LipidProbe/DataSets/8-3_FA_Huh7_SF_2024.csv") |>
# 	select(LipidProbe, gene_name, logFC, pvalue, hit_annotation)

# FA_1_10_data <- read_csv("LipidProbe/DataSets/1-10_FA_Huh7_SF_2024.csv") |>
# 	select(LipidProbe, gene_name, logFC, pvalue, hit_annotation)

# combinedDataset <- rbind(PA_data, PE_data, Spa_data, Sph_data, FA_8_3_data, FA_1_10_data)

#write.csv(combinedDataset, "LipidProbe/DataSets/combinedProbeDatasets_TMT.csv")
```

:::{.border}
:::{#fig-ShinyEmbed}
```{=html}
<!DOCTYPE html>
<html lang="en">
<head>
  <meta charset="UTF-8">
  <meta name="viewport" content="width=device-width, initial-scale=1.0">
  <title>Shiny App Embedding</title>
  <style>
    iframe {
      width: 100%;
      height: 90vh;
      border: none;
    }
  </style>
</head>
<body>
  <iframe src="https://gaelen-guzman.shinyapps.io/lipidinteractomics_shinyapp/"></iframe>
</body>
</html>
```

Black proteins are unenriched in respect to both probes; Purple proteins are "enriched" with respect to only the probe on the x-axis; Green proteins are "enriched" with respect to only the probe on the y-axis; and Orange proteins are "enriched" in both probes. Depicted are only the proteins present in the datasets of both probes.
:::
:::