### 2025-01-30 IndividualStudies/SF_2024.qmd

---
title: "Farley et al. 2024"
subtitle: "Trifunctional Sphinganine: A New Tool to Dissect Sphingolipid Function"
format: 
  html:
    embed-resources: false
standalone: false
---

```{r setup, echo=FALSE, include=FALSE}
## intitialize libraries
library(dplyr)
library(tidyr)
library(plotly)
library(ggplot2)
library(grid)
library(readr)
library(rmarkdown)
library(emojifont)
library(gt)
library(crosstalk)
library(DT)
library(htmltools)
library(scales)

## Opening the source file for the ggplot custom theme
source("ggplot_styles.R")

# Opening the datatable for this page IN FINAL RENDER
SF_2024_data <- readr::read_csv("DataTables/SF_ACS-Chem-Bio_2024_download.csv")

# # # Opening the datatable for this page WHILE EDITING
# SF_2024_data <- readr::read_csv("IndividualStudies/DataTables/SF_ACS-Chem-Bio_2024_download.csv")

# Organize the hit annotations as factors
SF_2024_data$hit_annotation <- factor(pull(SF_2024_data, hit_annotation), levels = c("enriched hit", "hit", "enriched candidate", "candidate", "no hit", "NA"), ordered = TRUE)

# Organize the probes as factors
SF_2024_data$LipidProbe <- factor(pull(SF_2024_data, LipidProbe), levels = c("Sph", "Spa", "8-3 FA", "1-10 FA"), ordered = TRUE)

# Building an ID column to maintain the order on the x-axis
ID <- c()
for(i in 0:(nrow(SF_2024_data)/2)){
	ID[i] = i
}

# Applies the ID column to the Spa results
SF_2024_ranked_Spa <- SF_2024_data |>
    arrange(logFC) |>
	filter(LipidProbe == "Spa") |>
	mutate(ID = ID)

# Applies the ID column to the Sph results
SF_2024_ranked_Sph <- SF_2024_data |>
    arrange(logFC) |>
	filter(LipidProbe == "Sph") |>
	mutate(ID = ID)

# Assembles the dataset to plot
SF_2024_ranked <- rbind(SF_2024_ranked_Sph, SF_2024_ranked_Spa)
```

::: {.callout-note collapse="true"}

{{< include /StudyInformation_includes/SF_2024_A_include.qmd >}}

:::

## Data visualization

::: {.panel-tabset}

#### Volcano plots

:::{.border}
```{r, echo=FALSE, warning=FALSE}
#| fig-cap: "Volcano plots depict the relative enrichment of each protein versus control (x-axis) and the statistical significance of each protein following a Student's t-test (y-axis, -log10 transformed). The left panel depicts the results of a pulldown using trifunctionalized phosphatidic acid; the right panel depicts the results of the pulldown using trifunctionalized phosphatidylethanolamine. Black proteins are unenriched or depleted in the presence of probe, Purple proteins are enriched to the probe but fall below a FDR or p-value threshold of significance, and Orange proteins surpass the threshold of significance (need details of threshold)."

VolcanoPlotStandardized(SF_2024_data)

```
:::

#### Ranked-order plots

:::{.border}
```{r, echo=FALSE, warning=FALSE}
#| fig-cap: "Ranked-order plots depict the relative enrichment of each protein versus control (y-axis) from lowest to highest. The left panel depicts the results of a pulldown using trifunctionalized phosphatidic acid; the right panel depicts the results of the pulldown using trifunctionalized phosphatidylethanolamine. Black proteins are unenriched or depleted in the presence of probe, Purple proteins are enriched to the probe but fall below a FDR or p-value threshold of significance, and Orange proteins surpass the threshold of significance (need details of threshold)."

RankedOrderPlotStandard(SF_2024_ranked)
```
:::

#### MA Plots

:::{.border}
```{r, echo=FALSE, warning=FALSE}
#| fig-cap: "MA plots"

MAStandard(SF_2024_data)

```

:::

:::

## Enriched Proteins

Check the boxes below to filter the dataset by which lipid probe was used in the pulldown and by significance thresholds.

:::{.border}
```{r, echo=FALSE, warning=FALSE}

# Here is the code to make the filterable lists shown in the crosstalk table below

SF_2024_data <- SF_2024_data |>
  select(gene_name, LipidProbe, hit_annotation, logFC, pvalue, fdr) |>
  mutate(logFC = signif(logFC, digits=3),
          pvalue = signif(pvalue, digits=3),
          fdr = signif(fdr, digits = 3)) |>
  arrange(hit_annotation) |>
  arrange(LipidProbe)

## Creates a reactive dataframe
SF_2024_data_reactive <- crosstalk::SharedData$new(SF_2024_data)

htmltools::tags$div(
  class = "filter-container",
  filter_checkbox(id = "LipidProbe",
                  label = "Lipid probe in pulldown:", 
                  sharedData = SF_2024_data_reactive,
                  group = ~LipidProbe),
  filter_checkbox(id = "hit_annotation",
                  label = "Enrichment criteria:", 
                  sharedData = SF_2024_data_reactive,
                  group = ~hit_annotation)
)

## Displays the reactive dataframe
datatable(SF_2024_data_reactive, 
          options = list(pageLength = 15),
                         colnames = c("Gene Name", "Lipid Probe", "Hit Annotation", "LogFC", "p value", "FDR"))

```
:::

## Data download

{{< downloadthis DataTables/SF_ACS-Chem-Bio_2024_download.csv dname="SF_ACS-Chem-Bio_2024">}}