### 2025-01-31 Background/DataAnalysisAndStatistics.qmd
Reorganizing sections, increasing visibility of subsections

---
title: "Data Analysis and Statistics"
---

## TMT quantitative mass spec pipeline

* Studies included in this repository which used identical FRAG Pipe TMT pipeline:
	- Thomas et al., 2025[@thomasTrifunctionalLipidDerivatives2025]
	- Farley et al., 2024 (A)[@farleyTrifunctionalSphinganineNew2024a]
	- Farley et al., 2024 (B)[@farleyTrifunctionalFattyAcid2024]

## PSM quantitative mass spec analysis

* Studies included in this repository which used a PSM-based analysis pipeline
	- Hoglinger et al., 2017[@hoglingerTrifunctionalLipidProbes2017]
	- Muller et al., 2021[@mullerSynthesisCellularLabeling2021]