### 2025-01-31 ContactUs/about.qmd

---
title: "About"
---

This repository is still under construction (as you can probably tell). Someday soon this page will have details about the founders of this resource, as well as a data submission form. Until then, please contact the site administrators for additional information about protocols, experimental design, and website feedback:

 <contact.us@LipidInteractome.org>{target="_blank"}

<br>
<hr>

### Github repository

This site was written by Gaelen Guzman and was built using Quarto. Find out more about Quarto [here](https://quarto.org){target="_blank"}.

This site can be pulled from GitHub at: <https://github.com/gaelenDG/LipidInteractomics_Website>

### Site structure

::::{.callout-note collapse="true"}

<!-- Run this to get a tree of the file structure, need to do some research on how this is depicted or if this is necessary -->
```{r setup, include=FALSE, echo=FALSE}
library(fs)

tree <- fs::dir_tree(path = ".", recurse = TRUE, depth = 3) |>
  stringr::str_subset("(^\\.|_site/|\\.git|\\.Rproj|\\.quarto|\\.cache)", negate = TRUE)

```

_site                                                
	├── 404.html                                         
	├── Background                                       
	│   ├── BackgroundFigures                            
	│   │   └── BaselineProteomics_ExperimentalDesign.png
	│   ├── BaselineProteomics_ExperimentalDesign.png    
	│   ├── DataAnalysisAndStatistics.html               
	│   ├── MultifunctionalLipidProbesOverview.html      
	│   └── ProteomicsUsingMultifunctionalProbes.html                
	├── ContactUs                                        
	│   ├── DataSubmission.html                          
	│   └── about.html                                   
	├── IndividualStudies                                
	│   ├── AT_2025.html                                 
	│   ├── DH_2017.html                                 
	│   ├── RM_2021.html                                 
	│   ├── SF_2024.html                                 
	│   ├── SF_2024_B.html                               
	│   └── StudyOverview.html                           
	├── LipidProbe                                       
	│   ├── 1-10_FattyAcid.html                          
	│   ├── 8-3_FattyAcid.html                           
	│   ├── Diacylglycerol.html                          
	│   ├── EnrichedHitsComparison.html                  
	│   ├── FormattingTemplate.html                      
	│   ├── PhosphatidicAcid.html                        
	│   ├── Phosphatidylethanolamine.html                
	│   ├── PhosphatidylinositolBisphosphate.html        
	│   ├── PhosphatidylinositolTrisphosphate.html       
	│   ├── Sphinganine.html                             
	│   ├── Sphingosine.html                             
	│   └── Structures                                   
	│       ├── 1-10_FattyAcid_Structure.png             
	│       ├── 8-3_FattyAcid_Structure.png              
	│       ├── PIP3_Structure.png                       
	│       ├── PhosphatidicAcid_Structure.png           
	│       ├── PhosphatidylethanolamineStructure.png    
	│       ├── SphinganineStructure.png                 
	│       └── SphingosineStructure.png                 
	├── MaintenanceInstructions.html                     
	├── Resources                                        
	│   └── styles.css                                   
	├── index.html

::::