### 2025-02-12 Background/DataAnalysisAndStatistics.qmd
updated data analysis background page

---
title: "Data Analysis and Statistics"
---

## TMT quantitative mass spec pipeline

Dr. Frank Stein developed and implemented the data analysis pipeline at the European Molecular Biosciences Laboratories in Heidelberg, Germany. The mass spectrometry (MS) data were processed using IsobarQuant[@mateusThermalProteomeProfiling2016] and Mascot V.24 (Matrix Science), following the methodology previously described by Farley et al. (2024, A), Farley et al. (2024, B), and Thomas et al (2025)[@farleyTrifunctionalSphinganineNew2024a; @farleyTrifunctionalFattyAcid2024; @thomasTrifunctionalLipidDerivatives2025]. To ensure robust protein identification, the analysis utilized a reverse UniProt FASTA database that included common contaminants. Proteins were only included in the final analysis if they contained at least two unique peptides, each with a minimum length of seven amino acids, and met a stringent false discovery rate (FDR) threshold of 0.01.

The acquired MS spectra were searched for specific peptide modifications, including fixed modifications (carbamidomethylation of cysteine and TMT-16-plex labeling of lysine) and variable modifications (N-terminal acetylation, methionine oxidation, and N-terminal TMT-16-plex labeling). Mass error tolerances were set to 10 ppm for full-scan MS spectra and 0.02 Da for MS/MS spectra, with a maximum of two missed cleavage sites permitted. To correct for batch effects across experimental replicates, a variance stabilization normalization was applied to log~2~-transformed raw data[@huberVarianceStabilizationApplied2002]. Changes in protein abundance between experimental and control conditions were assessed using LIMMA analysis[@ritchieLimmaPowersDifferential2015].

For the protein-lipid probe interaction experiments presented here (+/- UV exposure), proteins were classified as "enriched hits” if they showed at least a log~2~ fold-change of 1 toward the +UV condition with a p-value < 0.05. Proteins enriched under +UV conditions but with p-values > 0.05 were labeled as “enriched candidates.”

## PSM quantitative mass spec analysis

* Studies included in this repository which used a PSM-based analysis pipeline
	- Hoglinger et al., 2017[@hoglingerTrifunctionalLipidProbes2017]
	- Muller et al., 2021[@mullerSynthesisCellularLabeling2021]

* Need these details.