### 2025-02-20 IndividualStudies/DH_2017.qmd
Spent all day learning how to make my heatmap look less bad

---
title: "Höglinger et al. 2017"
subtitle: "Trifunctional lipid probes for comprehensive studies of single lipid species in living cells"
format:
  html:
    embed-resources: false
standalone: false
---
<!-- R setup, reading data tables, wrangling data -->
```{r setup, echo=FALSE, include=FALSE}

## intitialize libraries
library(dplyr)
library(tidyr)
library(plotly)
library(ggplot2)
library(grid)
library(readr)
library(rmarkdown)
library(emojifont)
library(gt)
library(crosstalk)
library(DT)
library(stringr)
library(pheatmap)
library(here)
library(heatmaply)
library(ggvenn)
library(scales)

## Opening the source file for the ggplot custom theme
source(paste0(here(), "/Resources/ggplot_styles.R"))

DH_2017_data <- read.csv(paste0(here(),"/IndividualStudies/DataTables/DH_PNAS_2017_download.csv")) |>
  mutate(gene_name = str_extract(Description, "GN=([^\\s]+)") |> str_remove("GN=")) |>
  mutate(Species = str_extract(Description, "OS=([^=]+?)\\sGN=") |> str_remove("OS=") |> str_remove("GN=")) |>
  filter(Species == "Homo sapiens ") |>
  mutate(PSM_Sph = (PSM_Sph1 + PSM_Sph2) / 2,
         PSM_FA = (PSM_FA1 + PSM_FA2) / 2,
         PSM_DAG = (PSM_DAG1 + PSM_DAG2) /2) |>
  select(gene_name, PSM_Sph, PSM_FA, PSM_DAG) |>
  arrange(gene_name) |>
  group_by(gene_name) |>
  summarise(PSM_DAG = mean(PSM_DAG),
            PSM_FA = mean(PSM_FA),
            PSM_Sph = mean(PSM_Sph))
```

::: {.callout-note collapse="true"}

{{< include /StudyInformation_includes/DH_2017_include.qmd >}}

:::

## Data visualization

#### Heat map
:::{.border}
```{r, echo=FALSE, fig.height=10}

DH_2017_data_long <- DH_2017_data |>
  pivot_longer(cols = starts_with("PSM_"),
              names_to = "variable",
              names_prefix = "PSM_",
              values_to = "value") |>
  mutate(name = gene_name) |>
  select(name, variable, value)
  # mutate(OriginalValues = value) |>
  # mutate(value = case_when(value < 20 ~ value,
  #                           value >= 20 ~ 20))

DH_2017_data_long$variable <- factor(pull(DH_2017_data_long, variable), ordered = TRUE, levels = c("Sph", "FA", "DAG"))

# Create the heatmap
heatmaply(long_data = DH_2017_data_long,
          xlab = "Probe",
          ylab = "Gene name",
          label_names = c("Gene name", "Probe", "PSM count"),
          scale_fill_gradient_fun = ggplot2::scale_fill_gradientn(colours = c("grey95", "purple", "orange", "orange", "orange", "orange", "orange", "red", "red","red"), breaks = c(0, 5, 20, 200)),
          showticklabels = F)

```
:::

<hr>

#### Venn Diagram

:::{.border}
```{r, echo=FALSE, warning=FALSE, height = 5}

# Ensure the data is in the correct format
DH_2017_data <- as.data.frame(DH_2017_data)
rownames(DH_2017_data) <- DH_2017_data[,1]
DH_2017_data[,1] <- NULL

Sph_proteins <- DH_2017_data |>
  filter(PSM_Sph > 0) |>
  select(PSM_Sph)

FA_proteins <- DH_2017_data |>
  filter(PSM_FA > 0) |>
  select(PSM_FA)

DAG_proteins <- DH_2017_data |>
  filter(PSM_DAG > 0) |>
  select(PSM_DAG)

combined_venn_proteins <- list("Sph proteins" = rownames(Sph_proteins), "FA proteins" = rownames(FA_proteins), "DAG proteins" = rownames(DAG_proteins))

ggvenn(combined_venn_proteins)

```
:::

## Data download

{{< downloadthis DataTables/DH_PNAS_2017_download.csv dname="DH_PNAS_2017">}}