### 2025-02-20 LipidProbe/Diacylglycerol.qmd

---
title: "Trifunctional Diacylglycerol"
prefer-html: true
---

## Structure

:::{.border}
:::

## Protein interactome

### Cell background: HeLa

::: {.callout-note collapse="true"}

{{< include /StudyInformation_includes/DH_2017_include.qmd >}}

:::