### 2025-02-23 IndividualStudies/AT_2025.qmd
Pre-rendering GO analysis so rendering is faster

---
title: "Thomas et al. 2025"
subtitle: "Trifunctional lipid derivatives: PE’s mitochondrial interactome"
format: 
  html:
    embed-resources: false
standalone: false
---
<!-- Data overview
Two lipid probes, PA and PE for a single cell line
 -->

::: {.callout-note collapse="true"}

{{< include /StudyInformation_includes/AT_2025_include.qmd >}}

:::

## Data visualization

```{r setup, echo=FALSE, include=FALSE}

## intitialize libraries
library(dplyr)
library(tidyr)
library(plotly)
library(ggplot2)
library(grid)
library(readr)
library(rmarkdown)
library(emojifont)
library(gt)
library(crosstalk)
library(DT)
library(here)

## Opening the source file for the ggplot custom theme
source(paste0(here(), "/Resources/ggplot_styles.R"))

```

```{r, echo=FALSE, include=FALSE}

# Data wrangling to prepare Alix's data to the standardized inputs for my plotting functions.
# Read the data to a df

AT_2025_Long <- readr::read_csv(paste0(here(), "/IndividualStudies/DataTables/AT_ChemicalCommunications_2025_download.csv"))

# Organize the hit annotations as factors
AT_2025_Long$hit_annotation <- factor(pull(AT_2025_Long, hit_annotation), levels = c("enriched hit", "enriched candidate", "no hit", "NA"), ordered=TRUE)

# Building an ID column to maintain the order on the x-axis
ID <- c()
for(i in 0:(nrow(AT_2025_Long)/2)){
	ID[i] = i
}

# Applies the ID column to the PE results
AT_2025_ranked_PE <- AT_2025_Long |>
    arrange(logFC) |>
	filter(LipidProbe == "PE") |>
	mutate(ID = ID)

# Applies the ID column to the PA results
AT_2025_ranked_PA <- AT_2025_Long |>
    arrange(logFC) |>
	filter(LipidProbe == "PA") |>
	mutate(ID = ID)

# Assembles the dataset to plot
AT_2025_ranked <- rbind(AT_2025_ranked_PE, AT_2025_ranked_PA)

```

::: {.panel-tabset}

#### Volcano plots

:::{.border}
```{r, echo=FALSE, warning=FALSE}
#| fig-cap: "Volcano plots depict the relative enrichment of each protein versus non-crosslinked control (x-axis) and the statistical significance of each protein following a Student's t-test (y-axis, -log10 transformed). The left panel depicts the results of a pulldown using trifunctionalized phosphatidic acid; the right panel depicts the results of the pulldown using trifunctionalized phosphatidylethanolamine. Black proteins are unenriched or depleted in the presence of probe, Purple enriched candidates are defined as proteins with a false discovery rate less than 0.2 and a fold change of at least 1.5-fold, and Orange enriched hits are defined as proteins with a false discovery rate less than 0.05 and a fold change of at least 2-fold in the +UV over the -UV)."

VolcanoPlotStandardized(AT_2025_Long)

```
:::

#### Ranked-order plots

:::{.border}
```{r, echo=FALSE, warning=FALSE}
#| fig-cap: "Ranked-order plots depict the relative enrichment of each protein versus non-crosslinked control (y-axis) from lowest to highest. The left panel depicts the results of a pulldown using trifunctionalized phosphatidic acid; the right panel depicts the results of the pulldown using trifunctionalized phosphatidylethanolamine. Black proteins are unenriched or depleted in the presence of probe, Purple enriched candidates are defined as proteins with a false discovery rate less than 0.2 and a fold change of at least 1.5-fold, and Orange enriched hits are defined as proteins with a false discovery rate less than 0.05 and a fold change of at least 2-fold in the +UV over the -UV)."

RankedOrderPlotStandard(AT_2025_ranked)
```
:::

#### MA Plots

:::{.border}
```{r, echo=FALSE, warning=FALSE}
#| fig-cap: "MA plots depict the average abundance of each protein in the crosslinked and uncrosslinked conditions (x-axis) versus the log2-transformed fold-change between the crosslinked and uncrosslinked conditions (y-axis). Black proteins are unenriched or depleted in the presence of probe, Purple enriched candidates are defined as proteins with a false discovery rate less than 0.2 and a fold change of at least 1.5-fold, and Orange enriched hits are defined as proteins with a false discovery rate less than 0.05 and a fold change of at least 2-fold in the +UV over the -UV)."

print("Data not yet provided.")

```

:::
:::

## Gene Ontology enrichment

**In beta: GO analysis still under development**

::: {.panel-tabset}

#### Cell Compartment

:::{.border}
```{r, echo=FALSE, warning=FALSE}
# #### These functions make the plots to be displayed in this div - uncomment them to run them again
# CC_enrichment_plots(AT_2025_Long, "dot", "PE-PA_AT_2025_CC-DOTplot")
# CC_enrichment_plots(AT_2025_Long, "cnet", "PE-PA_AT_2025_CC-CNETplot")
```

<iframe src="/IndividualStudies/GO_plots/PE-PA_AT_2025_CC-DOTplot.html" width="100%" height="500px"></iframe>

![](/IndividualStudies/GO_plots/PE-PA_AT_2025_CC-CNETplot.png)

:::

#### Molecular Function

:::{.border}
```{r, echo=FALSE, warning=FALSE}
#### These functions make the plots to be displayed in this div - uncomment them to run them again
# MF_enrichment_plots(AT_2025_Long, "dot", "PE-PA_AT_2025_MF-DOTplot")
# MF_enrichment_plots(AT_2025_Long, "cnet", "PE-PA_AT_2025_MF-CNETplot")
```

<iframe src="/IndividualStudies/GO_plots/PE-PA_AT_2025_MF-DOTplot.html" width="100%" height="500px"></iframe>

![](/IndividualStudies/GO_plots/PE-PA_AT_2025_MF-CNETplot.png)

:::

#### Biological Process

:::{.border}
```{r, echo=FALSE, warning=FALSE}
#### These functions make the plots to be displayed in this div - uncomment them to run them again
# BP_enrichment_plots(AT_2025_Long, "dot", "PE-PA_AT_2025_BP-DOTplot")
# BP_enrichment_plots(AT_2025_Long, "cnet", "PE-PA_AT_2025_BP-CNETplot")
```

<iframe src="/IndividualStudies/GO_plots/PE-PA_AT_2025_BP-DOTplot.html" width="100%" height="500px"></iframe>

![](/IndividualStudies/GO_plots/PE-PA_AT_2025_BP-CNETplot.png)

:::

:::

## Enriched proteins

Check the boxes below to filter the dataset by which lipid probe was used in the pulldown and by significance thresholds.

:::{.border}
```{r, echo=FALSE}

# Here is the code to make the filterable lists shown in the crosstalk table below

## Creates a reactive dataframe
AT_2025_Long_reactive <- crosstalk::SharedData$new(AT_2025_Long)

htmltools::tags$div(
  class = "filter-container",
  filter_checkbox(id = "LipidProbe",
                  label = "Lipid probe in pulldown:", 
                  sharedData = AT_2025_Long_reactive,
                  group = ~LipidProbe),
  filter_checkbox(id = "hit_annotation",
                  label = "Enrichment criteria:", 
                  sharedData = AT_2025_Long_reactive,
                  group = ~hit_annotation)
)

?crosstalk::filter_checkbox()

## Displays the reactive dataframe
datatable(AT_2025_Long_reactive, options = list(pageLength = 20))

```
:::

## Data download

{{< downloadthis DataTables/AT_ChemicalCommunications_2025_download.csv dname="AT_ChemicalCommunications_2025">}}