### 2025-02-24 SearchByProtein/SearchByProtein.qmd

---
title: "Search enrichment by protein"
prefer-html: true
---

```{r setup, include=FALSE}
library(DT)
library(crosstalk)
library(tidyverse)
library(readr)
library(ggplot2)
library(plotly)
library(here)

## Opening the source file for the ggplot custom theme
source(paste0(here(), "/Resources/ggplot_styles.R"))

data <- read_csv(paste0(here(), "/SearchByProtein/combinedProbeDatasets_TMT.csv"))

datawhole <- data |>
	dplyr::select(gene_name, LipidProbe, hit_annotation, logFC, pvalue, fdr) |>
	mutate(logFC = signif(logFC, digits=3),
          pvalue = signif(pvalue, digits=3),
					Study = case_when(LipidProbe == "1-10 FA" ~ "Farley et al. 2024 B; DOI: 10.1021/acschembio.3c00554",
														LipidProbe == "8-3 FA" ~ "Farley et al. 2024 B; DOI: 10.1021/acschembio.3c00554",
														LipidProbe == "Sph" ~ "Farley et al. 2024 A; DOI: 10.1021/acschembio.3c00554",
														LipidProbe == "Spa" ~ "Farley et al. 2024 A; DOI: 10.1021/acschembio.3c00554",
														LipidProbe == "PE" ~ "Thomas et al. 2025; DOI: 10.1039/D4CC03599B",
														LipidProbe == "PA" ~ "Thomas et al. 2025; DOI: 10.1039/D4CC03599B",
														LipidProbe == "PA" ~ "Thomas et al. 2025; DOI: 10.1039/D4CC03599B",
														)) |>
  arrange(hit_annotation) |>
  arrange(LipidProbe) |>
	glimpse()

limits <- c(max(abs(datawhole$logFC), na.rm=TRUE) , max(-log10(datawhole$pvalue), na.rm=TRUE))

# datawhole$LipidProbe <- factor(pull(data, LipidProbe), levels = c("PA", "PE", "Sph", "Spa", "1-10 FA", "8-3 FA"), ordered = TRUE)

	## Creates a reactive dataframe
datawhole_reactive <- crosstalk::SharedData$new(datawhole)

```

**Use the search bar below to filter for a protein among the results of all studies which utilized TMT-based quantification**

```{r, echo=FALSE}
htmltools::tags$div(
  class = "filter-container",
  filter_select(id = "gene_name",
                  label = "Protein name:", 
                  sharedData = datawhole_reactive,
                  group = ~gene_name)
)

## Displays the reactive dataframe
datatable(datawhole_reactive, 
          options = list(pageLength = 6,
												 dom = "l<t>ipr"),
												 colnames = c("Gene Name", "Lipid Probe", "Hit Annotation", "logFC", "pvalue", "FDR", "Study Details"))

```

"No hit" proteins are unenriched or depleted in the presence of probe, "enriched candidates" are defined as proteins with a false discovery rate less than 0.2 and a fold change of at least 1.5-fold, and "enriched hits" are defined as proteins with a false discovery rate less than 0.05 and a fold change of at least 2-fold in the +UV over the -UV

## Data vis of selected proteins

**Still in beta: a plot that highlights the selected protein out of all the datasets on our site**
```{r, warning=FALSE, echo=FALSE, fig.height=10}

VolcanoPlotStandardized(datawhole_reactive)

```