### 2025-02-27 LipidProbe/1-10_FattyAcid.qmd

---
title: "Trifunctional 1-10 Fatty Acid"
prefer-html: true
---

:::{.content-hidden unless-format="html"}

*Greater context to the distinction between the 8-3 Fatty Acid and 1-10 Fatty Acid probes can be found in the [Diazirine Overview](/Background/MultifunctionalLipidProbesOverview.qmd#sec-DiazirinePlacement)*

## Structure

:::{.border}
![Trifunctional 1-10 Fatty Acid](/LipidProbe/Structures/TF_1-10_fatty_acid.png)
:::

## Protein interactome

### Cell background: Huh7

<!-- R setup, reading dataframes and minor data wrangling -->
```{r setup, echo=FALSE, include=FALSE}
## intitialize libraries
library(here)

## Opening the source file for the ggplot custom theme
source(paste0(here(), "/Resources/ggplot_styles.R"))

# Opening the datatable for this page
data <- readr::read_csv(paste0(here(), "/LipidProbe/DataSets/1-10_FA_Huh7_SF_2024.csv")) |>
	filter(LipidProbe == "1-10 FA")

# Organizing the hit annotations as factors
data$hit_annotation <- factor(pull(data, hit_annotation), levels = c("enriched hit", "hit", "enriched candidate", "candidate", "no hit", "NA"), ordered = TRUE)

```

::: {.callout-note collapse="true"}

{{< include /StudyInformation_includes/SF_2024_B_include.qmd >}}

:::

:::: {.panel-tabset}

##### Volcano plot

:::{.border}
```{r, echo=FALSE, warning=FALSE}
#| fig-cap: "Data as reported in [Farley et al., 2024(B)](/IndividualStudies/SF_2024_B.qmd). Black proteins are unenriched or depleted in the presence of probe, Purple enriched candidates are defined as proteins with a false discovery rate less than 0.2 and a fold change of at least 1.5-fold, and Orange enriched hits are defined as proteins with a false discovery rate less than 0.05 and a fold change of at least 2-fold in the +UV over the -UV)."

VolcanoPlotStandardized(data)
```
:::

##### Ranked order plot

:::{.border}
```{r, echo=FALSE, warning=FALSE}
#| fig-cap: "Data as reported in [Farley et al., 2024(B)](/IndividualStudies/SF_2024_B.qmd). Black proteins are unenriched or depleted in the presence of probe, Purple enriched candidates are defined as proteins with a false discovery rate less than 0.2 and a fold change of at least 1.5-fold, and Orange enriched hits are defined as proteins with a false discovery rate less than 0.05 and a fold change of at least 2-fold in the +UV over the -UV)."

RankedOrderPlotStandard(data)
```
:::

##### MA Plots

:::{.border}
```{r, echo=FALSE, warning=FALSE}
#| fig-cap: "Data as reported in [Farley et al., 2024(B)](/IndividualStudies/SF_2024_B.qmd). Black proteins are unenriched or depleted in the presence of probe, Purple enriched candidates are defined as proteins with a false discovery rate less than 0.2 and a fold change of at least 1.5-fold, and Orange enriched hits are defined as proteins with a false discovery rate less than 0.05 and a fold change of at least 2-fold in the +UV over the -UV)."

MAStandard(data)

```
:::
:::

## Gene Ontology Analysis

**In beta: GO analysis still under development**

[Click here for more information about Gene Ontology analysis.](/Background/DataAnalysisAndStatistics.qmd#gene-ontology-overview)

:::::{.border}
::::{#fig-PE_GOplots}
::: {.panel-tabset}

#### Cell Compartment

<iframe src="/LipidProbe/GO_plots/1-10FA_SF_2024_CC-DOTplot.html" width="100%" height="500px"></iframe>

(CC CNET plot failed in production)

#### Molecular Function

(No MF pathways identified among enriched hits and candidates.)

(MF CNET plot failed in production)

#### Biological Process

(No BP pathways identified among enriched hits and candidates.)

(BP CNET plot failed in production)
:::

GO Dot plots display the enrichment of GO terms among the proteins enriched to the probe. For these analyses, only proteins categorized as "enriched candidates" and "enriched hits" were subject to GO analysis. The Cell Compartment analysis assesses whether the list of enriched proteins contains a statistically significant number of proteins in the same cellular region; appropriately, the Molecular Function and Biological Process analyses does the same for molecular function and biological process, respectively. [Click here for more information about GO dot plots.](/Background/DataAnalysisAndStatistics.qmd#gene-ontology-overview) 
::::
:::::

## Data exploration and download

:::{.border}
```{r, echo=FALSE, warning=FALSE}

# Here is the code to make the filterable lists shown in the crosstalk table below

data <- data |>
  dplyr::select(gene_name, LipidProbe, hit_annotation, logFC, pvalue, fdr) |>
  mutate(logFC = signif(logFC, digits=3),
          pvalue = signif(pvalue, digits=3),
          fdr = signif(fdr, digits = 3)) |>
  arrange(hit_annotation) |>
  arrange(LipidProbe)

## Creates a reactive dataframe
data_reactive <- crosstalk::SharedData$new(data)

htmltools::tags$div(
  class = "filter-container",
  filter_checkbox(id = "hit_annotation",
                  label = "Enrichment criteria:", 
                  sharedData = data_reactive,
                  group = ~hit_annotation,
									inline=TRUE)
)

## Displays the reactive dataframe
datatable(data_reactive, 
          caption = "'No hit' proteins are unenriched or depleted in the presence of probe, 'enriched candidates' are defined as proteins with a false discovery rate less than 0.2 and a fold change of at least 1.5-fold, and 'enriched hits' are defined as proteins with a false discovery rate less than 0.05 and a fold change of at least 2-fold in the +UV over the -UV", 
          options = list(pageLength = 10),
                         colnames = c("Gene Name", "Lipid Probe", "Hit Annotation", "LogFC", "p value", "FDR"))

```
:::

{{< downloadthis DataSets/1-10_FA_Huh7_SF_2024.csv dname="1-10_FA_Huh7_SF_ACS-Chem-Bio_2024">}}

::::

:::