### 2025-02-27 LipidProbe/Diacylglycerol.qmd
trying to distinguish cell types more effectively on individual probe pages

---
title: "Trifunctional Diacylglycerol"
prefer-html: true
callout-icon: false
---

## Structure

:::{.border}
:::

## Protein interactome

::::{.callout-tip}
### Cell background: HeLa

::: {.callout-note collapse="true"}

{{< include /StudyInformation_includes/DH_2017_include.qmd >}}

:::

::::