### 2025-03-04 LipidProbe/FormattingTemplate.qmd

---
title: "Formatting Template"
subtitle: "HTML page of individual lipid probe data"
---

## Structure

:::{.border}
![]()
:::

## Protein interactome

### Cell background: ???

<!-- Author information likely assembled in the relevant study page as well -->
:::: {.callout-note collapse="true"}

{{< include /StudyInformation_includes/template_include.qmd >}}

::::

<!-- Below is the default organization of the tabset with the standard Per/Frank TMT16 ratiometric data structure -->

::::{.panel-tabset}

<!-- R setup, reading dataframes and minor data wrangling -->
```{r setup, echo=FALSE, include=FALSE}
## intitialize libraries
library(dplyr)
library(tidyr)
library(plotly)
library(ggplot2)
library(grid)
library(readr)
library(rmarkdown)
library(emojifont)
library(gt)
library(crosstalk)
library(DT)
library(htmltools)
library(here)

## Opening the source file for the ggplot custom theme
source(paste0(here(), "/Resources/ggplot_styles.R"))

# # Opening the datatable for this page
# data <- readr::read_csv("/path/to/data")

# # Organizing the hit annotations as factors
# data$hit_annotation <- factor(pull(data, hit_annotation), levels = c("enriched hit", "hit", "enriched candidate", "candidate", "no hit", "NA"), ordered = TRUE)

```

##### Volcano plot

:::{.border}
```{r, echo=FALSE, warning=FALSE}
#| fig-cap: ""

# VolcanoPlotStandardized(data)
```
:::

##### Ranked order plot

:::{.border}
```{r, echo=FALSE, warning=FALSE}
#| fig-cap: ""

#RankedOrderPlotStandard(data)
```
:::

##### MA plot

:::{.border}
```{r, echo=FALSE, warning=FALSE}
#| fig-cap: ""

#MAStandard(data)

```
:::

##### Data

:::{.border}
```{r, echo=FALSE, warning=FALSE}

# Here is the code to make the filterable lists shown in the crosstalk table below

# data <- data |>
#   select(gene_name, LipidProbe, hit_annotation, logFC, pvalue, fdr) |>
#   mutate(logFC = signif(logFC, digits=3),
#           pvalue = signif(pvalue, digits=3),
#           fdr = signif(fdr, digits = 3)) |>
#   arrange(hit_annotation) |>
#   arrange(LipidProbe)

## Creates a reactive dataframe
# data_reactive <- crosstalk::SharedData$new(data)

# htmltools::tags$div(
#   class = "filter-container",
#   filter_checkbox(id = "hit_annotation",
#                   label = "Enrichment criteria:", 
#                   sharedData = data_reactive,
#                   group = ~hit_annotation,
# 									inline=TRUE)
# )

# ## Displays the reactive dataframe
# datatable(data_reactive, 
#           options = list(pageLength = 10),
#                          colnames = c("Gene Name", "Lipid Probe", "Hit Annotation", "LogFC", "p value", "FDR"))

```
:::

{{< downloadthis /path/to/file dname="filename.csv">}}

::::

<!-- And here is the default organization of a heatmap-based visualization -->

### Cell background: ???

<!-- Author information likely assembled in the relevant study page as well -->

:::: {.callout-note collapse="true"}

#### Study details

##### Authors

##### Journal

##### Abstract

##### Lipid probes utilized

##### Cell line analyzed

##### Mass spectrometry quantification method

::::

::::{.panel-tabset}

<!-- R chunk, reading dataframes and minor data wrangling -->
```{r, echo=FALSE, include=FALSE}

```

##### Heatmap
:::{.border}
```{r, echo=FALSE, warning=FALSE}
#| fig-cap: ""

#RankedOrderPlotStandard(data)
```
:::

##### Data

:::{.border}
```{r, echo=FALSE, warning=FALSE}

# Here is the code to make the filterable lists shown in the crosstalk table below

# data <- data |>
#   select(gene_name, LipidProbe, hit_annotation, logFC, pvalue, fdr) |>
#   mutate(logFC = signif(logFC, digits=3),
#           pvalue = signif(pvalue, digits=3),
#           fdr = signif(fdr, digits = 3)) |>
#   arrange(hit_annotation) |>
#   arrange(LipidProbe)

## Creates a reactive dataframe
# data_reactive <- crosstalk::SharedData$new(data)

# htmltools::tags$div(
#   class = "filter-container",
#   filter_checkbox(id = "hit_annotation",
#                   label = "Enrichment criteria:", 
#                   sharedData = data_reactive,
#                   group = ~hit_annotation,
# 									inline=TRUE)
# )

# ## Displays the reactive dataframe
# datatable(data_reactive, 
#           options = list(pageLength = 10),
#                          colnames = c("Gene Name", "Lipid Probe", "Hit Annotation", "LogFC", "p value", "FDR"))

```
:::

::::