### 2025-03-12 LipidProbe/PhosphatidylinositolTrisphosphate.qmd
updating structures

---
title: "Trifunctional Phosphatidylinositol 3,4,5-*tris*phosphate"
subtitle: "PI(3,4,5)P~3~"
prefer-html: true
---

## Structure

:::{.border}
::: {layout="[[-1.5,98.5], [100]]"}
![Native PI(3,4,5)P~3~.](/LipidProbe/Structures/PI(3,4)P2.png)

![Trifunctional PI(3,4,5)P~3~.](/LipidProbe/Structures/TF_PI(3,4,5)P3.png)

Chemical structures drawn by Berit Blume.
:::
:::

## Protein interactome