### 2025-03-25 IndividualStudies/DH_2017.qmd

---
title: "Höglinger et al. 2017"
subtitle: "Trifunctional lipid probes for comprehensive studies of single lipid species in living cells"
date: 2017-02-14
format:
  html:
    embed-resources: false
standalone: false
---
<!-- R setup, reading data tables, wrangling data -->
```{r setup, echo=FALSE, include=FALSE, warning=FALSE}

## intitialize libraries
library(here)

## Opening the source file for the ggplot custom theme
source(paste0(here(), "/Resources/ggplot_styles.R"))

###### Wrangling the raw dataset ######
DH_2017_data <- read.csv(paste0(here(),"/IndividualStudies/DataTables/DH_PNAS_2017_download.csv")) |>
  mutate(gene_name = str_extract(Description, "GN=([^\\s]+)") |> str_remove("GN=")) |>
  mutate(Species = str_extract(Description, "OS=([^=]+?)\\sGN=") |> str_remove("OS=") |> str_remove("GN=")) |>
  filter(Species == "Homo sapiens ") |>
  mutate(PSM_Sph = (PSM_Sph1 + PSM_Sph2) / 2,
         PSM_FA = (PSM_FA1 + PSM_FA2) / 2,
         PSM_DAG = (PSM_DAG1 + PSM_DAG2) /2) |>
  dplyr::select(gene_name, PSM_Sph, PSM_FA, PSM_DAG) |>
  arrange(gene_name) |>
  group_by(gene_name) |>
  summarise(PSM_DAG = mean(PSM_DAG),
            PSM_FA = mean(PSM_FA),
            PSM_Sph = mean(PSM_Sph))

###### Wrangling for the DataTable and the GO ######

DH_2017_data <- DH_2017_data |>
  mutate(VennRegion = case_when(
    (PSM_Sph > 0 & PSM_DAG > 0 & PSM_FA > 0) ~ "Enriched to all probes",
    (PSM_Sph > 0 & PSM_DAG == 0 & PSM_FA == 0) ~ "Enriched to Sph only",
    (PSM_Sph == 0 & PSM_DAG > 0 & PSM_FA == 0) ~ "Enriched to DAG only",
    (PSM_Sph == 0 & PSM_DAG == 0 & PSM_FA > 0) ~ "Enriched to FA only",
    (PSM_Sph > 0 & PSM_DAG > 0 & PSM_FA == 0) ~ "Enriched to Sph and DAG",
    (PSM_Sph > 0 & PSM_DAG == 0 & PSM_FA > 0) ~ "Enriched to Sph and FA",
    (PSM_Sph == 0 & PSM_DAG > 0 & PSM_FA > 0) ~ "Enriched to DAG and FA"))

DH_2017_data$VennRegion <- factor(pull(DH_2017_data, VennRegion), ordered = TRUE, levels = c("Enriched to Sph only", "Enriched to DAG only", "Enriched to FA only", "Enriched to Sph and DAG", "Enriched to Sph and FA", "Enriched to DAG and FA", "Enriched to all probes"))

###### Wrangling for the HeatMap ######

DH_2017_data_HeatMap <- DH_2017_data |>
  pivot_longer(cols = starts_with("PSM_"),
              names_to = "variable",
              names_prefix = "PSM_",
              values_to = "value") |>
  mutate(name = gene_name) |>
  dplyr::select(name, variable, value)

DH_2017_data_HeatMap$variable <- factor(pull(DH_2017_data_HeatMap, variable), ordered = TRUE, levels = c("Sph", "FA", "DAG"))

###### Wrangling for the Venn Diagram ######

# Separates out each of the genesets, then forces them back together in a vectorized list.
Sph_proteins <- DH_2017_data |>
  filter(PSM_Sph > 0) |>
  dplyr::select(gene_name)

FA_proteins <- DH_2017_data |>
  filter(PSM_FA > 0) |>
  dplyr::select(gene_name)

DAG_proteins <- DH_2017_data |>
  filter(PSM_DAG > 0) |>
  dplyr::select(gene_name)

DH_2017_VennData <- list("Sph proteins" = as.list(Sph_proteins$gene_name), "FA proteins" = as.list(FA_proteins$gene_name), "DAG proteins" = as.list(DAG_proteins$gene_name))

```

::: {.callout-note collapse="true"}

{{< include /StudyInformation_includes/DH_2017_include.qmd >}}

:::

## Data visualization

#### Heat map

:::{.border}

```{r Making Heat Map, echo=FALSE, fig.height=10, warning=FALSE}
#| fig-cap: "Heat map depicts the average PSM count for two replicates each of UV-irradiated samples treated with trifunctionalized Sph, DAG, and FA. Color denotes the "

# Create the heatmap
heatmaply(long_data = DH_2017_data_HeatMap,
          xlab = "Probe",
          ylab = "Gene name",
          label_names = c("Gene name", "Probe", "PSM count"),
          scale_fill_gradient_fun = ggplot2::scale_fill_gradientn(colours = c("grey95", "purple", "orange", "orange", "orange", "orange", "orange", "red", "red","red"), breaks = c(0, 5, 20, 200)),
          showticklabels = F)

```
:::

<hr>

#### Venn Diagram

:::{.border}

```{r Making Venn Diagram, echo=FALSE, warning=FALSE, width = 8}
#| fig-cap: "A Venn diagram depicts the overlapping identification of proteins between the three lipid probes. There are 41 proteins uniquely identified after enrichment with tf-Sph, 130 proteins uniquely identified after enrichment with tf-DAG, and no proteins uniquely identified to tf-8:3-FA."

ggvenn <- ggvenn(DH_2017_VennData,
                      show_percentage = FALSE,
                      show_elements = FALSE,
                      fill_color = c("orange", "purple", "blue")) +
        labs(x = "", y = "") +
        theme_minimal()

ggvenn

```

:::

## Gene Ontology Analysis

**In beta: GO analysis still under development**

:::::{.border}
::::{#fig-PE_GOplots}
::: {.panel-tabset}

#### Cell Compartment

<iframe src="/IndividualStudies/GO_plots/DH_2017_CC_DOTPLOT.html" width="100%" height="500px"></iframe>

#### Molecular Function

<iframe src="/IndividualStudies/GO_plots/DH_2017_MF_DOTPLOT.html" width="100%" height="500px"></iframe>

#### Biological Process

<iframe src="/IndividualStudies/GO_plots/DH_2017_BP_DOTPLOT.html" width="100%" height="500px"></iframe>

:::

GO Dot plots display the enrichment of GO terms among the proteins enriched to the probe. For these analyses, only proteins categorized as "enriched candidates" and "enriched hits" were subject to GO analysis. The Cell Compartment analysis assesses whether the list of enriched proteins contains a statistically significant number of proteins in the same cellular region; appropriately, the Molecular Function and Biological Process analyses does the same for molecular function and biological process, respectively. [Click here for more information about Gene Ontology analysis.](/Background/DataAnalysisAndStatistics.qmd#gene-ontology-overview)
::::
:::::

## Data exploration

:::{.border}
```{r Building Datatable, echo=FALSE, warning=FALSE}
DH_2017_data_reactive <- crosstalk::SharedData$new(DH_2017_data)

htmltools::tags$div(
  class = "filter-container",
  filter_checkbox(id = "Overlap",
                  label = "Select overlap regions to see co-enriched proteins:",
                  sharedData = DH_2017_data_reactive,
                  group = ~VennRegion,
                  inline = TRUE)
)

## Displays the reactive dataframe
datatable(DH_2017_data_reactive, options = list(pageLength = 20))
```

:::

## Data download

{{< downloadthis DataTables/DH_PNAS_2017_download.csv dname="DH_PNAS_2017">}}