### 2025-03-25 IndividualStudies/RM_2021.qmd

---
title: "Muller et al. 2021"
subtitle: "Synthesis and Cellular Labeling of Multifunctional Phosphatidylinositol *Bis*- and *Tris*phosphate Derivatives"
date: 2021-06-01
format: 
  html:
    embed-resources: false
standalone: false
---

```{r setup, echo=FALSE, include=FALSE}
## intitialize libraries
library(here)

## Opening the source file for the ggplot custom theme
source(paste0(here(), "/Resources/ggplot_styles.R"))

# Opening the datatable for this page WHILE TESTING
# RM_2021_data <- readr::read_csv("IndividualStudies/DataTables/RM_Angewandte_2021_download.csv")

# Opening the datatable for this page WHILE RENDERING
RM_2021_data <- readr::read_csv("DataTables/RM_Angewandte_2021_download.csv")

```

::: {.callout-note collapse="true"}

{{< include /StudyInformation_includes/RM_2021_include.qmd >}}

:::

## Data visualization

#### Heat map

#### Venn Diagram

## Data download

{{< downloadthis DataTables/RM_Angewandte_2021_download.csv dname="RM_Angewandte_2021">}}