### 2025-03-28 IndividualStudies/WY_2021.qmd
adding WY_2021 study, checking if authors like it

---
title: "Yu et al. 2021"
subtitle: "A Chemoproteomics Approach to Profile Phospholipase D-Derived Phosphatidyl Alcohol Interactions"
date: 2021-12-15
format: 
  html:
    embed-resources: false
standalone: false
---

::: {.callout-note collapse="true"}

{{< include /StudyInformation_includes/WY_2021_include.qmd >}}

:::

```{r NewPlottingFunctions, include=FALSE, echo=FALSE}
RankedOrderPlotSpecial <- function(data) {
    ################
    #' A standardized Ranked Order plot theme
    #' Requires an input dataframe with the columns "gene_name", "LipidProbe", "logFC", "pvalue", and "hit_annotation"
    ################

    # Applies ID column to each probe
    data <- data |>
      group_by(LipidProbe) |>
      arrange(logFC, .by_group = TRUE) |>
      mutate(ID = row_number()) |>
      ungroup()

    # First makes a ggplot template to later apply ggplotly
    RankedOrderPlots <- data |>
        ggplot(aes(x = ID,
                   y = logFC,
                   color = factor(hit_annotation),
                   fill = factor(hit_annotation),
                   shape = factor(hit_annotation),
                   alpha = factor(hit_annotation),
                   size = factor(hit_annotation))) +
        geom_point(aes(text = paste0("Gene name: ", gene_name, "\n",
                                    "LogFC: ", logFC, "\n"))) +
        geom_hline(yintercept = 0, linetype = 2) +
        geom_vline(xintercept = 0, linetype = 2) +
        scale_shape_manual(values = c("enriched hit" = 21,
                                    "hit" = 21,
                                    "enriched candidate" = 24,
                                    "candidate" = 24,
                                    "no hit" = 22), 
                        name = "Trend") +
        scale_color_manual(values = c("enriched hit" = "black",
                                    "hit" = "black",
                                    "enriched candidate" = "black",
                                    "candidate" = "black",
                                    "no hit" = "black"), 
                        name = "Trend") +
        scale_fill_manual(values = c("enriched hit" = "orange",
                                    "hit" = "orange",
                                    "enriched candidate" = "purple",
                                    "candidate" = "purple",
                                    "no hit" = "black"), 
                        name = "Trend") +
        scale_alpha_manual(values = c("enriched hit" = 0.8,
                                    "hit" = 0.8,
                                    "enriched candidate" = 0.5,
                                    "candidate" = 0.5,
                                    "no hit" = 0.25), 
                        name = "Trend") +
        scale_size_manual(values = c("enriched hit" = 4,
                                    "hit" = 4,
                                    "enriched candidate" = 2,
                                    "candidate" = 2, 
                                    "no hit" = 0.75), 
                        name = "Trend") +
        facet_wrap(~LipidProbe) +
        ylab("Log2 fold-change") +
        xlab("Ranked-order gene") +
        theme_minimal() +
        theme(text = element_text(family = "serif")) +
        theme(panel.border = element_rect(color = NA, fill = NA)) +
        theme(legend.title = element_text(size = 12, face = "bold")) +
        theme(axis.title = element_text(size = 12, face = "bold")) +
        theme(legend.text = element_text(size = 12))  +
        theme(strip.text = element_text(
    size = 20))
    m <- list(l=50, r = 50, b= 100, pad = 4)

    RankedOrderPlotly <- ggplotly(RankedOrderPlots,
                            tooltip = c("text", dynamicTicks = TRUE),
                            hovermode = "closest",
                            autosize = T,
                            margin = m) |>
                    layout(showlegend = FALSE)

    return(RankedOrderPlotly)
}

MASpecial <- function(data) {

    ################
    #' A standardized MA plot function
    #' Receives a dataframe. Needs to have the following columns for the aes() to work: AveExpr, logFC, hit_annotation, gene_name.
    #' Dynamically determines the limits of the plot under the scale_x/y_scontinuous calls
    #' Produces first a facet-wrapped ggplot, then converts it to a plotly plot.
    ################

    MAPlots <- data |>
        ggplot(aes(x = AveExpr,
                   y = logFC,
                   color = factor(hit_annotation),
                   fill = factor(hit_annotation),
                   shape = factor(hit_annotation),
                   alpha = factor(hit_annotation),
                   size = factor(hit_annotation))) +
        geom_point(aes(text = paste0("Gene name: ", gene_name, "\n",
                                    "LogFC: ", logFC, "\n",
                                    "AveExpr: ", AveExpr, "\n"))) +
        geom_hline(yintercept = 0, linetype = 2) +
        geom_vline(xintercept = 0, linetype = 2) +
        scale_x_continuous(limits = c(0.9 * min(abs(data$AveExpr)), 1.1 * max(abs(data$AveExpr))))+
        scale_y_continuous(limits = c(-1.1 * max(abs(data$logFC)), 1.1 * max(abs(data$logFC))))+
        scale_shape_manual(values = c("enriched hit" = 21,
                                    "hit" = 21,
                                    "enriched candidate" = 24,
                                    "candidate" = 24,
                                    "no hit" = 22), 
                        name = "Trend") +
        scale_color_manual(values = c("enriched hit" = "black",
                                    "hit" = "black",
                                    "enriched candidate" = "black",
                                    "candidate" = "black",
                                    "no hit" = "black"), 
                        name = "Trend") +
        scale_fill_manual(values = c("enriched hit" = "orange",
                                    "hit" = "orange",
                                    "enriched candidate" = "purple",
                                    "candidate" = "purple",
                                    "no hit" = "black"), 
                        name = "Trend") +
        scale_alpha_manual(values = c("enriched hit" = 0.8,
                                    "hit" = 0.8,
                                    "enriched candidate" = 0.5,
                                    "candidate" = 0.5,
                                    "no hit" = 0.25), 
                        name = "Trend") +
        scale_size_manual(values = c("enriched hit" = 4,
                                    "hit" = 4,
                                    "enriched candidate" = 2,
                                    "candidate" = 2, 
                                    "no hit" = 0.75), 
                        name = "Trend") +
        facet_wrap(~LipidProbe) +
        ylab("Log2 fold-change") +
        xlab("Average Ion Intensity") +
        theme_minimal() +
        theme(text = element_text(family = "serif")) +
        theme(panel.border = element_rect(color = NA, fill = NA)) +
        theme(legend.title = element_text(size = 12, face = "bold")) +
        theme(axis.title = element_text(size = 12, face = "bold")) +
        theme(legend.text = element_text(size = 12)) +
        theme(strip.text = element_text(
    size = 20))

    m <- list(l=50, r = 50, b= 100, pad = 4)

    MAPlotly <- ggplotly(MAPlots,
                            tooltip = c("text", dynamicTicks = TRUE),
                            hovermode = "closest",
                            autosize = T,
                            margin = m) |>
                    layout(showlegend = FALSE)

    return(MAPlotly)
}

```

```{r Setup_Data_Wrangling, include=FALSE, echo=FALSE}
## intitialize libraries
library(here)
library(biomaRt)
library(rentrez)

## Opening the source file for the ggplot custom theme
source(paste0(here(), "/Resources/ggplot_styles.R"))

# Opening raw data file for wrangling and cleaning up
RawData <- read_csv(paste0(here::here(), "/IndividualStudies/DataTables/WY_2021_ACSChemBio_RawData.csv"))

#RawData <- read_csv("IndividualStudies/DataTables/WY_2021_ACSChemBio_RawData.csv")

RawData <- RawData |>
  filter(Coverage != 0 & Fold_change != 0) |>
  group_by(Accession) |>
  summarise(gene_name = Description,
            logFC = mean(log2(Fold_change)),
            # pvalue = mean(pvalue),
            numReplicates = n(),
            AveExpr = log10(mean(Control_abundance) + mean(Sample_abundance)/2)) |>
  arrange(-logFC)

median_logFC <- median(RawData$logFC)

RawData$logFC <- RawData$logFC - median_logFC

RawData <- RawData |> 
  mutate(LipidProbe = "PDAA",
        AveExpr = case_when(.default = AveExpr,
                              AveExpr == -Inf ~ 0),
                              hit_annotation = case_when(.default = "no hit",
                                                          logFC >= 0.5849625 & numReplicates == 2 ~ "enriched candidate",
                                                          logFC >= 0.5849625 & numReplicates == 3 ~ "enriched hit")) |>         
  glimpse()

nrow(RawData |> filter(hit_annotation == "enriched hit"))

# Organize the hit annotations as factors
RawData$hit_annotation <- factor(pull(RawData, hit_annotation), levels = c("enriched hit", "hit", "enriched candidate", "candidate", "no hit", "NA"), ordered = TRUE)

# write_csv(RawData, paste0(here(), "/IndividualStudies/DataTables/WY_2021_ACSChemBio_Download.csv"))

```

## Data visualization

::::{.border}
::: {.panel-tabset}

##### Ranked order plot

```{r, echo=FALSE, warning=FALSE}
#| fig-cap: "Data as reported in [Yu et al., 2021, 2024](/IndividualStudies/WY_2021.qmd). Black proteins are identified in only one replicate, Purple enriched candidates are defined as proteins identified in two replicates & fold-change > 1.5 (logFC 0.585) & pvalue < 0.01, and Orange enriched hits are defined as proteins identified in all three replicates & fold-change > 1.5 (logFC 0.585) & pvalue < 0.01)."

RankedOrderPlotSpecial(RawData)
```

##### MA plot

```{r, echo=FALSE, warning=FALSE}
#| fig-cap: "Data as reported in [Yu et al., 2021](/IndividualStudies/WY_2021.qmd). Black proteins are identified in only one replicate, Purple enriched candidates are defined as proteins identified in two replicates & fold-change > 1.5 (logFC 0.585) & pvalue < 0.01, and Orange enriched hits are defined as proteins identified in all three replicates & fold-change > 1.5 (logFC 0.585) & pvalue < 0.01)."

MASpecial(RawData)
```

##### Venn Diagram

```{r Venn_wrangling, include=FALSE, warning=FALSE}

###### Wrangling for the Venn Diagram ######
# Read the dataset

VennData <- read_csv("DataTables/WY_2021_ACSChemBio_RawData.csv")  |>
  mutate(gene_name = Description)

# Separates out each of the genesets, then forces them back together in a vectorized list.
A_proteins <- VennData |>
  filter(Trial == 1) |>
  dplyr::select(gene_name)

B_proteins <- VennData |>
  filter(Trial == 2) |>
  dplyr::select(gene_name)

C_proteins <- VennData |>
  filter(Trial == 3) |>
  dplyr::select(gene_name)

VennList <- list("Trial 1" = as.list(A_proteins$gene_name), "Trial 2" = as.list(B_proteins$gene_name), "Trial 3" = as.list(C_proteins$gene_name))
```

```{r NAPE_Venn, echo=FALSE, warning=FALSE}
#| fig-cap: "Data as reported in [Yu et al., 2021](/IndividualStudies/WY_2021.qmd). A Venn diagram depicts the overlapping identification of proteins between the three trials (replicates)."

ggvenn <- ggvenn(VennList,
                      show_percentage = FALSE,
                      show_elements = FALSE,
                      fill_color = c("orange", "purple", "blue")) +
        labs(x = "", y = "") +
        theme_minimal()

ggvenn

```

:::
::::

## Data exploration

:::{.border}
```{r NAPE_DataTable, echo=FALSE, warning=FALSE}

# Here is the code to make the filterable lists shown in the crosstalk table below

data <- RawData |>
  dplyr::select(Accession, gene_name, LipidProbe, hit_annotation, logFC, numReplicates) |>
  mutate(logFC = signif(logFC, digits=3)) |>
  arrange(-logFC) |>
  arrange(hit_annotation)

## Creates a reactive dataframe
data_reactive <- crosstalk::SharedData$new(data)

htmltools::tags$div(
  class = "filter-container",
  filter_checkbox(id = "LipidProbe",
                  label = "Lipid probe in pulldown:", 
                  sharedData = data_reactive,
                  group = ~LipidProbe),
  filter_checkbox(id = "hit_annotation",
                  label = "Enrichment criteria:", 
                  sharedData = data_reactive,
                  group = ~hit_annotation)
)

## Displays the reactive dataframe
datatable(data_reactive, 
          caption = "This datatable contains all the data plotted above. Check the boxes below to filter the dataset by which lipid probe was used in the pulldown and by significance thresholds. logFC ratios are median normalized. \"Num Replicate IDs\" denotes the number of replicates in which the protein was identified. As above, \"enriched hit\" denotes proteins identified in all three replicates & fold-change > 1.5 (logFC 0.585) & pvalue < 0.01. \"enriched candidates\" are proteins identified in two replicates & fold-change > 1.5 (logFC 0.585) & pvalue < 0.01.",
          options = list(pageLength = 15),
                         colnames = c("Accession", "Gene Name", "Lipid Probe", "Hit Annotation", "LogFC", "Num Replicate IDs"),
                         rownames=FALSE)

```

:::

<br>

## Data download

{{< downloadthis DataTables/WY_2021_ACSChemBio_Download.csv dname="WY_2021_ACS-ChemBio_Download">}}