### 2025-03-28 StudyInformation_includes/DCC_2025_include.qmd

## Study details

#### Authors

Din-Chi Chiu, Hening Lin, Jeremy M. Baskin

#### Journal
ChemR&Chi;iv

<https://doi.org/10.26434/chemrxiv-2025-j8nqz>{target="_blank"}

#### Abstract
Most eukaryotic membranes comprise phospholipids bearing two hydrophobic tails, but N-acylphosphatidylethanolamine (NAPE) stands out as a long-known but poorly understood phospholipid with three hydrophobic groups. What little attention that NAPE has received has been devoted to understanding its metabolic functions as a precursor to N-acylethanolamines, bioactive lipids that act as endocannabinoids. Yet, levels of NAPE increase during myocardial infarction and ischemia, suggesting potential signaling roles for this lipid. Here, we exploit photoaffinity labeling (PAL) to identify NAPE-interacting proteins and elucidate signaling functions of NAPE. By positioning diazirine and alkyne groups in metabolically distinct regions of the NAPE molecule, we ensured that our PAL probe reported on interactions of NAPE and not NAE. Our studies identified several NAPE interactors, including two single-pass transmembrane proteins, CD147/Basigin and CD44, both of which serve as chaperones for monocarboxylate transporters (MCTs) from the SLC16A family that mediate lactate flux across the plasma membrane. Functional studies revealed that NAPE stimulates lactate efflux by MCTs dependent upon CD147 and CD44, establishing NAPE as a bona fide signaling lipid and pointing to potential physiological roles in metabolic and energy homeostasis that may be pathologically relevant in ischemia.

#### Lipid probes utilized

<a href="/LipidProbe/N-Acylphosphatidylethanolamine.qmd" class="hover-preview">Bifunctional NAPE</a>
<span class="hover-image">
  <img src="/LipidProbe/Structures/BF_NAPE.png" alt="Bifunctional NAPE" />
</span>

#### Cell line analyzed

[HeLa](https://www.atcc.org/products/ccl-2){target="_blank"}

#### Uncaging & Crosslinking timeline

| Lipid Probe | Uptake time | Interaction time | Crosslinking time |
|:-------|:------:|:------:|:------:|:------:|
| LEI-401    | 30 min  | 60 min | 30 min  |

#### Mass spectrometry quantification method

SILAC

#### Additional sample preparation <span class="help-symbol" title="e.g. sub-cellular fractionation, depletion, etc.">?</span>