### 2025-06-16 HourlyTracking/202505_GG-Hours.qmd
Updating with data wrangling scripts and some suggested reviewer changes

---
title: "April 2025 Hours Report"
author: "Gaelen Guzman"
date: "2025 05 02"
output: 
  pdf_document:
    keep_tex: true
fontsize: 12pt
geometry: margin=1in
header-includes:
  - "\\usepackage{fancyhdr}"
  - "\\pagestyle{fancy}"
  - "\\fancyhead[L]{Gaelen Guzman}"
  - "\\fancyhead[C]{Hours Report}"
  - "\\fancyhead[R]{2025 06 02}"
  - "\\fancyfoot{}"
  - "\\renewcommand{\\headrulewidth}{0.4pt}"
  - "\\renewcommand{\\footrulewidth}{0pt}"
---

Carsten Schultz,

Please find below the hours I logged for the month of May 2025, with descriptions of each day's activity. In total, I performed 8 hours of work at an hourly rate of $50 (USD), for a total bill of $400.

Please forward payments to the following account:

| | |
| :--- | :--- |
| Bank | Wells Fargo | 
| Account name |  Gaelen Guzman | 
| Address | 2925 SE Clinton St., Portland, OR 97202 |
| Account number | 8583230647 | 
| Routing number | 124103799 |

```{r, echo=FALSE, warning=FALSE, include = FALSE}
# filepath: [Hours.csv](http://_vscodecontentref_/0)
library(readr)
library(kableExtra)
library(knitr)
library(dplyr)
library(here)

# Read the Hours.csv file and handle unnamed columns
hours_data <- read_csv(here("HourlyTracking/hours.csv"), col_names = c("Date", "Hours", "Activity"), show_col_types = FALSE) |>
  mutate(Date = as.Date(as.character(Date), format = "%Y%m%d")) |>
  select(Date, Hours, Activity) |>
  filter(Date >= as.Date("2025-05-01") & Date < as.Date("2025-06-01")) %>%  # Filter for April 2025
  glimpse()

sum(hours_data$Hours)

```

```{r, echo=FALSE}

# Display the table with text wrapping for the "Activity" column
kable(
  hours_data,
  format = "latex",
  booktabs = TRUE,
  caption = "Hours Logged",
  label = "hours-logged"
) %>%
  column_spec(3, width = "10cm")  # Adjust the width of the "Activity" column
```

<br>

Thank you,

Gaelen Guzman