### 2025-07-29 LipidProbe/Diacylglycerol.qmd
Spent a long time learning about custom divs, applied a new one for cell types

---
title: "Trifunctional Diacylglycerol"
format: 
  html:
    embed-resources: false
---

## Structure

:::{.border}
![(A) Trifunctional Diacylglycerol. Preliminary structure collected from Höglinger et al. 2017](/LipidProbe/Structures/TF_DAG.png)
:::

## Protein interactome

::::{.callout-tip}

### Cell background: HeLa

::: {.callout-note collapse="true"}

{{< include /StudyInformation_includes/DH_2017_include.qmd >}}

:::

::::