### 2025-07-29 LipidProbe/PhosphatidylinositolTrisphosphate.qmd

---
title: "Trifunctional Phosphatidylinositol 3,4,5-*tris*phosphate"
subtitle: "PI(3,4,5)P~3~"
format: 
  html:
    embed-resources: false
---

## Structure

:::{.border}
::: {layout="[[-1.5,98.5], [100]]"}
![(A) Native PI(3,4,5)P~3~.](/LipidProbe/Structures/PI(3,4)P2.png)

![(B) Trifunctional PI(3,4,5)P~3~.](/LipidProbe/Structures/TF_PI(3,4,5)P3.png)

Chemical structures drawn by Berit Blume.
:::
:::

## Protein interactome