### 2025-07-29 Resources/testing.qmd

::: todo
Remember to complete this section.
:::

::: test
This information is crucial for understanding the concept.
:::

::: {.thumbs-up title="That was a hard task!"}
You did a great job completing this task.
:::

::: {.cells}
Does this do it??
:::

:::{.cells}
::::{.cells-header}
Membrane fraction
::::
::::{.cells-container}
* Know how to make your own callouts.
* Be able to mess with some SCSS/CSS styling.
::::
:::

::: {.callout-tip}

## Membrane Fraction

* content
* content
* content
:::