### 2025-08-29 LipidProbe/PhosphatidylinositolTrisphosphate.qmd

---
title: "Trifunctional Phosphatidylinositol 3,4,5-*tris*phosphate"
subtitle: "PI(3,4,5)P~3~"
format: 
  html:
    embed-resources: false
---

## Structure

:::{.border}
::: {layout="[[-1.5,98.5], [100]]"}
![(A) Native PI(3,4,5)P~3~.](/LipidProbe/Structures/PI(3,4)P2.png)

![(B) Trifunctional PI(3,4,5)P~3~.](/LipidProbe/Structures/TF_PI(3,4,5)P3.png)

Chemical structures drawn by Berit Blume.
:::
:::

## Protein Interactome

```{r setup, echo=FALSE, include=FALSE}
## intitialize libraries
library(here)
here()
## Opening the source file for the ggplot custom theme
source((here("Resources/ggplot_styles.R")))

# Opening the datatable for this page
RM_2021_data <- readr::read_csv(here("LipidProbe/DataSets/PIP3_HeLa_RM_2021_download.csv")) |>
  mutate(
    hit_annotation = case_when(
      hit_annotation == "enriched hit" & logFC < 0 ~ "hit",
      hit_annotation == "enriched candidate" & logFC < 0 ~ "candidate",
      .default = hit_annotation
    )
  )

# Organize the hit annotations as factors
RM_2021_data$hit_annotation <- factor(pull(RM_2021_data, hit_annotation), levels = c("enriched hit", "hit", "enriched candidate", "candidate", "no hit",  "NA"), ordered=TRUE)

RM_2021_data <- RM_2021_data |>
  arrange(hit_annotation) |>
  glimpse()
```

::: {.callout-note collapse="true"}

{{< include /StudyInformation_includes/RM_2021_include.qmd >}}

:::

::: {.callout-tip}

### Membrane Fraction

```{r FilteringForMembraneProteins, include=FALSE, echo=FALSE}
data <- RM_2021_data |>
  filter(Fraction == "Membrane (+UV vs -UV)")

```

#### Data visualization

::: {.panel-tabset}

##### Volcano plots

:::{.border}
```{r, echo=FALSE, warning=FALSE}
#| fig-cap: "Volcano plots depict the relative enrichment of each protein versus non-crosslinked control (x-axis) and the statistical significance of each protein following a Student's t-test (y-axis, -log10 transformed). Original publication reported results in terms of peptide spectral matches; here, data have been reanalyzed to compare normalized intensities of experimental (+UV)and control (-UV) samples without isobaric mass tags. Black proteins are unenriched or depleted in the presence of probe, Purple enriched candidates are defined as proteins with a false discovery rate less than 0.2 and a fold change of at least 1.5-fold, and Orange enriched hits are defined as proteins with a false discovery rate less than 0.05 and a fold change of at least 2-fold in the +UV over the -UV)."

VolcanoPlotStandardized(data)

```
:::

##### Ranked-order plots

:::{.border}
```{r, echo=FALSE, warning=FALSE}
#| fig-cap: "Ranked-order plots depict the relative enrichment of each protein versus non-crosslinked control (y-axis) from lowest to highest. Original publication reported results in terms of peptide spectral matches; here, data have been reanalyzed to compare normalized intensities of experimental (+UV)and control (-UV) samples without isobaric mass tags. Black proteins are unenriched or depleted in the presence of probe, Purple enriched candidates are defined as proteins with a false discovery rate less than 0.2 and a fold change of at least 1.5-fold, and Orange enriched hits are defined as proteins with a false discovery rate less than 0.05 and a fold change of at least 2-fold in the +UV over the -UV)."

RankedOrderPlotStandard(data)
```
:::

##### MA Plots

:::{.border}
```{r, echo=FALSE, warning=FALSE}
#| fig-cap: "MA plots depict the average abundance of each protein in the crosslinked and uncrosslinked conditions (x-axis) versus the log2-transformed fold-change between the crosslinked and uncrosslinked conditions (y-axis). Original publication reported results in terms of peptide spectral matches; here, data have been reanalyzed to compare normalized intensities of experimental (+UV)and control (-UV) samples without isobaric mass tags. Black proteins are unenriched or depleted in the presence of probe, Purple enriched candidates are defined as proteins with a false discovery rate less than 0.2 and a fold change of at least 1.5-fold, and Orange enriched hits are defined as proteins with a false discovery rate less than 0.05 and a fold change of at least 2-fold in the +UV over the -UV)."

MAStandard(data)

```

:::
:::
:::

::: {.callout-tip}

### Cytosolic Fraction

```{r FilteringForCytosolProteins, include=FALSE, echo=FALSE}
data <- RM_2021_data |>
  filter(Fraction == "Cytosol (+UV vs -UV)")

```

#### Data visualization

::: {.panel-tabset}

##### Volcano plots

:::{.border}
```{r, echo=FALSE, warning=FALSE}
#| fig-cap: "Volcano plots depict the relative enrichment of each protein versus non-crosslinked control (x-axis) and the statistical significance of each protein following a Student's t-test (y-axis, -log10 transformed). Original publication reported results in terms of peptide spectral matches; here, data have been reanalyzed to compare normalized intensities of experimental (+UV)and control (-UV) samples without isobaric mass tags. Black proteins are unenriched or depleted in the presence of probe, Purple enriched candidates are defined as proteins with a false discovery rate less than 0.2 and a fold change of at least 1.5-fold, and Orange enriched hits are defined as proteins with a false discovery rate less than 0.05 and a fold change of at least 2-fold in the +UV over the -UV)."

VolcanoPlotStandardized(data)

```
:::

##### Ranked-order plots

:::{.border}
```{r, echo=FALSE, warning=FALSE}
#| fig-cap: "Ranked-order plots depict the relative enrichment of each protein versus non-crosslinked control (y-axis) from lowest to highest. Original publication reported results in terms of peptide spectral matches; here, data have been reanalyzed to compare normalized intensities of experimental (+UV)and control (-UV) samples without isobaric mass tags. Black proteins are unenriched or depleted in the presence of probe, Purple enriched candidates are defined as proteins with a false discovery rate less than 0.2 and a fold change of at least 1.5-fold, and Orange enriched hits are defined as proteins with a false discovery rate less than 0.05 and a fold change of at least 2-fold in the +UV over the -UV)."

RankedOrderPlotStandard(data)
```
:::

##### MA Plots

:::{.border}
```{r, echo=FALSE, warning=FALSE}
#| fig-cap: "MA plots depict the average abundance of each protein in the crosslinked and uncrosslinked conditions (x-axis) versus the log2-transformed fold-change between the crosslinked and uncrosslinked conditions (y-axis). Original publication reported results in terms of peptide spectral matches; here, data have been reanalyzed to compare normalized intensities of experimental (+UV)and control (-UV) samples without isobaric mass tags. Black proteins are unenriched or depleted in the presence of probe, Purple enriched candidates are defined as proteins with a false discovery rate less than 0.2 and a fold change of at least 1.5-fold, and Orange enriched hits are defined as proteins with a false discovery rate less than 0.05 and a fold change of at least 2-fold in the +UV over the -UV)."

MAStandard(data)

```

:::
:::
:::

## Data exploration

Check the boxes below to filter the dataset by which lipid probe was used in the pulldown and by significance thresholds.

:::{.border}
```{r, echo=FALSE}

RM_2021_data <- RM_2021_data |>
  select(-hit_annotation_method, -Protein.Description) |>
  mutate(
    logFC = signif(logFC, digits = 4),
    pvalue = signif(pvalue, digits = 4),
    AveExpr = signif(AveExpr, digits = 4),
    fdr = signif(fdr, digits = 4)
  )

# Here is the code to make the filterable lists shown in the crosstalk table below

## Creates a reactive dataframe
data_reactive <- crosstalk::SharedData$new(RM_2021_data)

htmltools::tags$div(
  class = "filter-container",
  filter_checkbox(id = "LipidProbe",
                  label = "Lipid probe in pulldown:", 
                  sharedData = data_reactive,
                  group = ~LipidProbe),
  filter_checkbox(id = "hit_annotation",
                  label = "Enrichment criteria:", 
                  sharedData = data_reactive,
                  group = ~hit_annotation),
  filter_checkbox(id = "Fraction",
                  label = "Cellular fraction",
                  sharedData = data_reactive,
                  group = ~Fraction)
)

## Displays the reactive dataframe
datatable(data_reactive, 
          options = list(pageLength = 10),
          colnames = c("Rank", "Gene Name", "Accession #","Lipid Probe", "Fraction", "LogFC", "p value", "FDR", "Hit Annotation"))

```

NB: Data analysis overview pending.
:::

## Data download

{{< downloadthis DataSets/PIP3_HeLa_RM_2021_download.csv dname="PIP3_HeLa_RM_2021">}}