### 2025-08-29 StudyInformation_includes/AT_2025_include.qmd

## Study details

#### Authors
Alix Thomas, Rainer Mueller, Ana Kojic, Scotland E. Farley, Frank Stein, Per Haberkant, Carsten Schultz

#### Journal
Chemical Communications

<https://doi.org/10.1039/D4CC03599B>{target="_blank"}

#### Abstract
Phosphatidylethanolamine (PE) is a ubiquitous lipid species in higher eukaryotes which resides preferentially in mitochondria. Here, we synthesized a multifunctionalized PE derivative (1) de-signed to identify PE-binding proteins in intact cells through photo-crosslinking and subsequent isolation and proteomic analysis of the PE-protein conjugates. Due to its aromatic caging group, 1 is initially concentrating in perinuclear membranes. After uncaging, rapid translocation to mitochondria was observed. Hence, the tool is useful for tracking PE location and for determining the PE interactome. A trifunctional phosphatidic acid (PA) derivative in compari-son was rapidly metabolized and is hence more limited in its use.

#### Lipid probes utilized

<a href="/LipidProbe/PhosphatidicAcid.qmd" class="hover-preview">Trifunctional Phosphatidic Acid</a>
<span class="hover-image">
  <img src="/LipidProbe/Structures/TF_PA.png" alt="Phosphatidic Acid" />
</span>

<a href="/LipidProbe/Phosphatidylethanolamine.qmd" class="hover-preview">Trifunctional Phosphatidylethanolamine</a>
<span class="hover-image">
  <img src="/LipidProbe/Structures/TF_PE.png" alt="Phosphatidylethanolamine" />
</span>

#### Cell line analyzed

[HeLa](https://www.atcc.org/products/ccl-2){target="_blank"}

#### Uncaging & Crosslinking timeline

| Lipid Probe | Uptake time | Uncaging time | Interaction time | Crosslinking time |
|:-------|:------:|:------:|:------:|:------:|
| PA    | 60 min | 5 min  | 15 min | 5 min  |
| PE    | 60 min | 5 min  | 15 min | 5 min  |

#### Mass spectrometry quantification method

16-channel Tandem Mass Tagging (TMT16)

::: {.callout-note collapse="true"}
##### ProteomeXchange details

**ProteomeXchange title:** Trifunctional lipid derivatives: PE’s mi-tochondrial interactome

**ProteomeXchange accession:** PXD059960

**PubMed ID:** Not applicable

**Publication DOI:** 10.1039/D4CC03599B

**Project Webpage:** http://www.ebi.ac.uk/pride/archive/projects/PXD059960

**FTP Download:** https://ftp.pride.ebi.ac.uk/pride/data/archive/2025/01/PXD059960
:::

#### Additional sample preparation <span class="help-symbol" title="e.g. sub-cellular fractionation, depletion, etc.">?</span>

Cell lysates were separated into membrane and cytosolic fractions. This publication reports the results of just the membrane fraction.

#### Data wrangling
Use the download button below to download the R script used to wrangle the authors' original submission into the data visualization tools on the Lipid Interactome:

{{< downloadthis /Users/gaelenguzman/LipidInteractomics_Website_local/LipidInteractomics_Website/IndividualStudies/DataTables/DataWrangling/AT_ChemComm2025_wrangling.R dname="AT_ChemComm2025_wrangling">}}