### 2025-08-29 StudyInformation_includes/RM_2021_include.qmd

## Study details

#### Authors
Rainer Muller, Ana Kojic, Mevlut Citir, Carsten Schultz

#### Journal
Angewandte Chemie

<https://doi.org/10.1002/anie.202103599>{target="_blank"}

#### Abstract

We synthesized the first multifunctionalized phosphoinositide polyphosphate derivatives featuring a photo-removable protecting group (“cage”), a photo-crosslinkable diazirine group, and a terminal alkyne group useful for click chemistry. We demonstrate that the lipid derivatives readily enter cells. After photo-crosslinking, cell fixation and fluorescent tagging via click chemistry, we determined the intracellular location of the lipid derivatives before and after uncaging of the lipids. We find that there is rapid trafficking of PI(3,4)P2 and PI(3,4,5)P3 derivatives to the plasma membrane, opening the intriguing possibility that there is active transport of these lipids involved. We employed the photo-crosslinking and click chemistry functions to analyze the proteome of PI(3,4,5)P3-binding proteins. From the latter, we validated by RNAi that the putative lipid binding proteins ATP11A and MPP6 are involved in the transport of PI(3,4,5)P3 to the plasma membrane.

#### Lipid probes utilized

[Trifunctional PI(3,4,5)P~3~](/LipidProbe/PhosphatidylinositolTrisphosphate.qmd)

#### Cell line analyzed

[HeLa](https://www.atcc.org/products/ccl-2){target="_blank"}

#### Mass spectrometry quantification method

Peptide spectral counting *(in original publication)*

Label free quantification *(as reported here, enabled via fresh analysis by Frank Stein, EMBL)*

#### Additional sample preparation <span class="help-symbol" title="e.g. sub-cellular fractionation, depletion, etc.">?</span>

Samples were split into membrane and cytosolic fractions and analyzed as independent mass spec samples. Additional *No Probe* controls were included for raw input, though not used in displaying the data on this site.

#### Data wrangling
Use the download button below to download the R script used to wrangle the authors' original submission into the data visualization tools on the Lipid Interactome:

{{< downloadthis /Users/gaelenguzman/LipidInteractomics_Website_local/LipidInteractomics_Website/IndividualStudies/DataTables/DataWrangling/RM_2021_Wrangling.R dname="RM_2021_wrangling">}}