### 2025-09-04 LipidProbe/Diacylglycerol.qmd
ok one last change, made the sph, 8-3 FA  and dag pages show a line about developing the PSM data visualization, and added collapsible panels

---
title: "Trifunctional Diacylglycerol"
format: 
  html:
    embed-resources: false
---

## Structure

:::{.border}
![(A) Trifunctional Diacylglycerol. Preliminary structure collected from Höglinger et al. 2017](/LipidProbe/Structures/TF_DAG.png)
:::

## Protein interactome

::::{.callout-tip}

### Cell background: HeLa

::: {.callout-note collapse="true"}

{{< include /StudyInformation_includes/DH_2017_include.qmd >}}

:::

**\*PSM data presentation still in development.**

See [Höglinger et al. 2017](https://lipidinteractome.org/individualstudies/dh_2017) for results in context with Diacylglycerol and 8-3 Fatty Acid data.

::::