### 2025-10-31 ContactUs/about.qmd

---
title: "About"
---

This repository is still under construction (as you can probably tell). Someday soon this page will have details about the founders of this resource, as well as a data submission form. Until then, please contact the site administrators for additional information about protocols, experimental design, and website feedback:

 <contact.us@LipidInteractome.org>{target="_blank"}

<br>
<hr>

### Github repository

This site was written by Gaelen Guzman and was built using Quarto. Find out more about Quarto [here](https://quarto.org){target="_blank"}.

This site can be pulled from GitHub at: <https://github.com/gaelenDG/LipidInteractomics_Website>